---
title: "Statistical Inference - Project B - Claudio Bogazzi"
output:
  pdf_document:
    latex_engine: xelatex
---

In this 2nd project we are going to analyze the ToothGrowth data in the R datasets package.
This data comes from an essay testing the eﬀect of Vitamin C on toothgrowth in Guinea Pigs. 
It consists of the measured tooth length in 6 groups of 10 animals, 
divided into two groups depending on the delivery method of the Vitamin C (OJ or VJ):
Each of the two groups is also divided into 3 subgroups according to the dosage, 0.5, 1 or 2 mg per dose

**1. Load the ToothGrowth data and perform some basic exploratory data analyses.**

```{r}

library(datasets)
data(ToothGrowth)
dim(ToothGrowth)

head(ToothGrowth)

```


**2. Provide a basic summary of the data.**
We immediately provide the summary of the data set using the R function summary().

```{r}
summary(ToothGrowth)
```
We now make two simple histograms showing the distribution length of the tooth for the two
different delivery methods.
```{r}
#a couple of simple histograms

lenvc <- ToothGrowth$len[which(ToothGrowth$supp == "VC")]
lenoj <- ToothGrowth$len[which(ToothGrowth$supp == "OJ")]

h <- hist(lenvc,xlab="Lenth of Tooth")

hh <- hist(lenoj, xlab ="Length of Tooth")

```

The next table shows the dose levels (rows) for the different delivery methods (columns)

```{r}

# table to show dose levels and delivery methods
table(ToothGrowth$dose, ToothGrowth$supp)

```
Finally, we plot the dose levels as a function of the length of the tooth for the two methods.

```{r}
#dose vs tooth lenth

library(ggplot2)
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity",) +
    facet_grid(. ~ supp) +
    xlab("Dose in miligrams") +
    ylab("length") +
    guides(fill=guide_legend(title="Supp type"))


```


**3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.**

We now compute the different confidence intervals by using the t.test() function in R.
First we consider the case where the dose is equal to 0.5. 
Let's see if there is a statistically significant
difference in the length of tooth associated with
the two different delivery methods.


```{r}

dose <- ToothGrowth$dose
supp <- ToothGrowth$supp
len <- ToothGrowth$len

data1 <- len[which(supp=="OJ" & dose==0.5)]
data2 <- len[which(supp=="VC" & dose==0.5)]

t.test(data1,data2)


```


Indeed there is a difference. Moreover, since both limits of the confidence intervals 
are positive we can say that the mean of the length when the delivery method is "OJ"
is greater than the alternative.

Let's now have a look at the case where the dose is equal to 1.0 and 2.0.

```{r}

data3 <- len[which(supp=="OJ" & dose==1.0)]
data4 <- len[which(supp=="VC" & dose==1.0)]

t.test(data3,data4)

data5 <- len[which(supp=="OJ" & dose==2.0)]
data6 <- len[which(supp=="VC" & dose==2.0)]

t.test(data5,data6)

```

The results obtained for the case dose = 1.0 are similar to dose = 0.5.
A different result is obtained for dose = 2.0 where we can see that the means
for the two delivery methods are similar. Indeed, the confidence interval
contains 0.

**4. State your conclusions and the assumptions needed for your conclusions.**

The mean value of the length of the tooth when the delivery method "OJ" is is greater than
when the delivery method is "VC" for both doses of 0.5 and 1.0.

When the dose is 2.0 there are no evidence that the means of the length of the tooth for
the two delivery methods is different.
